Estimating Neighborhood Rents with Scraped Data

### 2018-06-15 manuscript/index.Rmd
catching typos, working on the helper scripts some

Census data have common currency among scholars and practitioners interested in metropolitan housing markets, but neverthless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more. This publication lag prohibits analyzing market dynamics beyond long-term trends, particularly for smaller areas of aggregation like neighborhoods where the estimate may be an average of survey responses from a 5 year period. While more up-to-date options for neighborhood rent data exist through privately-held companies, such data can be somewhat of a black box in their creation and may be biased towards the costs of corporate-held properties. This would-be problem for academic researchers creates an opportunity for applying web scraping methods to generate a novel data on housing availability and costs.

@@

Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market. Both data suggest continued increases in the marketwide median rent for the first half of 2017, though the Craigslist slope is about twice in its size. Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for 2017 Q2 upward in favor of regularly posted commerical listings. Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability. This volatility in Craigslist asking rents could also be brought on by recently-developed buildings in "lease-up" adding a glut of expensive listings to the quarters' samples.

@@

Though the spatial distribution of the ACS estimates resembles that of the other two data sources, the values are nonethless systematically lower (in most cases by hundreds of dollars). The ACS measures therefore give insight into the relative ranking of neighborhoods in terms of rental costs, but are limited for studying neighborhood dynamics outside of long-run trend.

@@

Figure 5 shows a scatterplot/bubbpleplot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 75 tracts where comparable estimates exist. The purple dashed line indicates the loess smoother fit between the measures. Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and good agreement between the measures (mean absolute difference = $186). The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e. the asking rent less an incentives a landlord offers). Given the currently low vacancy rates in most properties, the only neighborhoods where there would theoretically be more salience to the asking vs contract rent distinction (i.e. a wider gap in these indicators) are at the high end of the distribution where demand for luxury apartments has slackened some. At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.

@@

An alternative explanation fpr the difference between measures comes in considering the units these properties are sampling from in a given neighborhood. The greater difference between the measures among lower-rent neighborhoods (with Craigslist estimating these as more expensive than Dupre+Scott) provides some evidence for this theory. These tracts are found towards the North and South edges of the city, and generally are more residential neighborhoods where 20+ unit apartment complexes are not the entire stock of rental unit properties. In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf){target="_blank"}.

@@

The BYM Spatial with AR(1) is the [Besags-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466){target="_blank"} intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand). The random-effect $S_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $S_i|S_j, j \in ne(i) ~ \mbox{Normal}(\bar{S_i}, \frac{\sigma^{2}_s}{m_j})$. Lastly, the BYM Spatiotemporal AR(1) (`spt`) adds an exchangeable random-effect for each period to allow for area-specific deviations in time trends.

@@

Finally, this paper has endeavored (perhaps a little too much...) to provide an example of using scraped rental housing listings for research, but also to model how Rmarkdown, Git and GitHub are transformative  for our disciplines. There are no free lunches with reproducibility: these open-source technologies demand that we learn how to incorporate them into our work. To the extent that this requires thinking about analysis beyond a userbase of 1, there can be some growing pains. However, there are some positive spillovers beyond improving the _science_ of social science---including new platforms and media for sharing research findings.

>>>>

Census data have common currency among scholars and practitioners interested in metropolitan housing markets, but nevertheless possess a substantial limitation for research in tight housing markets like Seattle---that is, at least two years elapse between data collection and publication, if not more. This publication lag prohibits analyzing market dynamics beyond long-term trends, particularly for smaller areas of aggregation like neighborhoods where the estimate may be an average of survey responses from a 5 year period. While more up-to-date options for neighborhood rent data exist through privately-held companies, such data can be somewhat of a black box in their creation and may be biased towards the costs of corporate-held properties. This would-be problem for academic researchers creates an opportunity for applying web scraping methods to generate a novel data on housing availability and costs.

@@

Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market. Both data suggest continued increases in the market-wide median rent for the first half of 2017, though the Craigslist slope is about twice in its size. Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for 2017 Q2 upward in favor of regularly posted commercial listings. Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability. This volatility in Craigslist asking rents could also be brought on by recently-developed buildings in "lease-up" adding a glut of expensive listings to the quarters' samples.

@@

Though the spatial distribution of the ACS estimates resembles that of the other two data sources, the values are nonetheless systematically lower (in most cases by hundreds of dollars). The ACS measures therefore give insight into the relative ranking of neighborhoods in terms of rental costs, but are limited for studying neighborhood dynamics outside of long-run trend.

@@

Figure 5 shows a scatter plot/bubble plot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 75 tracts where comparable estimates exist. The purple dashed line indicates the loess smoother fit between the measures. Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and good agreement between the measures (mean absolute difference = $186). The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e. the asking rent less an incentives a landlord offers). Given the currently low vacancy rates in most properties, the only neighborhoods where there would theoretically be more salience to the asking vs contract rent distinction (i.e. a wider gap in these indicators) are at the high end of the distribution where demand for luxury apartments has slackened some. At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.

@@

An alternative explanation for the difference between measures comes in considering the units these properties are sampling from in a given neighborhood. The greater difference between the measures among lower-rent neighborhoods (with Craigslist estimating these as more expensive than Dupre+Scott) provides some evidence for this theory. These tracts are found towards the North and South edges of the city, and generally are more residential neighborhoods where 20+ unit apartment complexes are not the entire stock of rental unit properties. In these tracts, Craigslist has a non-zero probability of sampling from any larger building but also small houses, duplex/triplexes, mother-in-law suites and recently, [Accessory Dwelling Units (ADUs)](http://www.seattle.gov/DPD/Publications/CAM/cam116a.pdf){target="_blank"}.

@@

The BYM Spatial with AR(1) is the [Besag-York-Mollie (1991)](https://link.springer.com/article/10.1007/BF00116466){target="_blank"} intrinsic conditional autoregressive spatial model with AR(1) and exchangeable hyperparameters for the panel's temporal structure (referred to as `bym` using shorthand). The random-effect $S_i$ is assumed to come from a normal distribution whose mean is the mean of the neighbors' random effects and variance proportional to the number of neighbors (formally, $S_i|S_j, j \in ne(i) ~ \mbox{Normal}(\bar{S_i}, \frac{\sigma^{2}_s}{m_j})$. Lastly, the BYM Spatiotemporal AR(1) (`spt`) adds an exchangeable random-effect for each period to allow for area-specific deviations in time trends.

@@

Finally, this paper has endeavored (perhaps a little too much...) to provide an example of using scraped rental housing listings for research, but also to model how Rmarkdown, Git and GitHub are transformative  for our disciplines. There are no free lunches with reproducibility: these open-source technologies demand that we learn how to incorporate them into our work. To the extent that this requires thinking about analysis beyond a user base of 1, there can be some growing pains. However, there are some positive spillovers beyond improving the _science_ of social science---including new platforms and media for sharing research findings.
